Estimating migration rates for iteroparous species using kinship pairs found between populations

### 2021-06-25 latex/ms.tex
write 1st para in 2.2, #6

\section{Theory}\label{sec2}
In this section, we present the theoretical foundation for estimating a migrant number and a migration rate using HS or PO pairs, which are found between two populations with a different sample timing. While the estimators can hold under flexible assumptions for reproduction (see {\bf Appendix}), here we assume a relatively simple situation for explanation purposes. It should be noted that the formulation of the estimators, presented in Equations 4-5 and 7-8, is a function with parent numbers for each population; in the next section, we provide the treatment for estimating the unknown parent numbers from genetic dataset which is used for obtaining HS and/or PO pairs. The main symbols used in the current paper are summarized in Table \ref{symbols}.

@@

Suppose that we have a set of two populations labelled 1 and 2 where random mating occurs within each population and that parents can move to the other population after the end of the reproductive season. Hereafter, without loss of generality, we consider the movement of parents in the direction from population 1 to population 2 and focus on the migration number or rate, and that is what we attempt to estimate. In this framework, considering two reproductive seasons (hereafter, we call them "the first year" and "the second year") is enough to estimate it.

Suppose there are $N_1$ parents in the population 1 at the beginning of the first year. Each parent produces offspring in which the number is determined by the parent's reproductive potential, denoted by $\lambda_{i,1}$ ($i=1,2,\ldots,N_1$). After the reproductive season, part of the parents would start to moving towards the population 2; $M$ survived migrants arrive in the population 2 so that there are $N_2$ parents in the population 2 at the beginning of the second year ($M<N_2$). Similar to the population 1, $N_2$ parents produce offspring in which the number is determined by $\lambda_{j,2}$ ($j=1,2,\ldots,N_2$).

@@

To estimate the migration number or rate, we utilize the number of HS and/or PO pairs observed in a sample. In the both populations, $n_{\rm P, 1}$ and $n_{\rm P, 2}$ parents are randomly sampled immediately after the end of the reproductive season in the first and second year, respectively. Additionally, $n_{\rm O, 1}$ and $n_{\rm O, 2}$ offspring are also randomly sampled. It should be noted that the four kinds of a sample are not always required for estimating migration number or rate; what kinds of samples are needed depends on the situation, as explained later.

@@

Figure \ref{cartoon} illustrates an example of kinship relationships and movements of parents between populations.

@@

\renewcommand{\arraystretch}{0.6}
\begin{table}[tb]
\begin{center}
   \caption[]{Summary of proposed estimators for required parameters and kinship types}
    \textbf {}\\[-0mm]
    \begin{tabular}{ccccccccc} \hline
       	\\
	Estimator			& $N_1$ 		& $N_2$ 		& $n_{\rm O, 1}$	& $n_{\rm O, 2}$	& $n_{\rm P, 1}$	& $n_{\rm P, 2}$	& Required kinship type\\
	\\
	\hline
	\\
	$\widehat{M_1}$	& given 		& given 		& \checkmark		& \checkmark 		&				&				& HSP\\ 
		                						& \\
	$\widehat{M_2}$	& given		& given		& \checkmark		& 				&				& \checkmark 		& POP\\
		                						& \\
	$\widehat{M_3}$	& given		& given		& \checkmark		&  \checkmark		&				& \checkmark 		& HSP \& POP\\
		                						& \\
	$\widehat{M_4}$	& given		& estimated	& \checkmark		&  \checkmark		&				& \checkmark 		& HSP \& POP\\
		                						& \\
	$\widehat{M_5}$	& estimated	& estimated	& \checkmark		&  \checkmark		& \checkmark		& \checkmark 		& HSP \& POP\\
									& \\
	$\widehat{m_1}$	& ---			& given		& \checkmark		& \checkmark 		&				&				& HSP\\ 
									& \\
	$\widehat{m_2}$	& ---			& given		& \checkmark		& 				&				& \checkmark 		& POP\\					                			& \\
	$\widehat{m_3}$	& ---			& given 		& \checkmark		&  \checkmark		&				& \checkmark 		& HSP \& POP\\
		                						& \\
	$\widehat{m_4}$	& ---			& estimated	& \checkmark		&  \checkmark		& 				& \checkmark 		& HSP \& POP\\
		                						& \\
	\hline
    \end{tabular}
    \label{parameter} 
\end{center} 
\end{table}
\renewcommand{\arraystretch}{1}

@@

\renewcommand{\arraystretch}{0.6}
\begin{table}[tb]
\begin{center}
   \caption[]{The list of mathematical symbols employed in the main text}
    \textbf {}\\[-4mm]
    \begin{tabular}{llc} \hline
       & & \\
	$n_{\rm P,1}, n_{\rm P, 2}$			& Sample number of parents in the population 1 and 2\\ 
		                						& \\
	$n_{\rm O, 1}, n_{\rm O, 2}$			& Sample number of offspring in the population 1 and 2\\ 
		                						& \\
	$N_{1}, N_{2}$						& Number of parents in the population 1 and 2 when sampled offspring are born\\
		                						& \\
	$M$								& Number of survived migrants of parents from the population 1 to population 2\\
		                						& \\
	$r$								& Sex ratio\\
		                						& \\
	$\pi_{\rm PO,1}, \pi_{\rm PO,2}$		& Probability that a randomly selected pair (mother and offspring) \\
	                							& shares a parent-offspring relationship within the population 1 and 2\\
									& \\
	$\pi_{\rm PO, bet}$					& Probability that a randomly selected pair (parent and offspring in the population 1 and 2, respectively) \\
	                							& shares a parent-offspring relationship between the population 1 and 2\\
									& \\
	$\pi_{\rm HS, bet}$					& Probability that a randomly selected pair (two offspring in the population 1 and 2) \\
	                							& shares a half-sibling relationship between the population 1 and 2\\
					                			& \\
	$\lambda_{i,1}, \lambda_{j,2}$			& Expected number of surviving offspring of parent $i$ and $j$ at sampling in the population 1 and 2\\
		                						& \\
	$\lambda_{i, M}$					& Expected number of surviving offspring of migrant $i$ at sampling\\
		                						& \\
	$k_{i,1}, k_{j,2}$					& Number of surviving offspring born to parent $i$ and $j$ in the population 1 and 2\\ 
	                							& \\
	$H_{\rm PO, 1}, H_{\rm PO, 2}$		& Number of parent-offspring pairs observed in samples within the population 1 and 2\\ 
	                							& \\
	$H_{\rm PO, bet}$					& Number of parent-offspring pairs observed in samples between the population 1 and 2\\ 
	                							& \\
	$H_{\rm HS, bet}$					& Number of half-sibling pairs observed in samples between the population 1 and 2\\ 
	                							& \\              		
	\hline
    \end{tabular}
    \label{symbols} 
\end{center} 
\end{table}
\renewcommand{\arraystretch}{1}

@@

@@

>>>>

\section{Theory}\label{sec2}
In this section, we present the theoretical foundation for estimating a migrant number and a migration rate using HS or PO pairs, which are found between two populations with a different sample timing. While the estimators can hold under flexible assumptions for reproduction (see {\bf Appendix}), here we assume a relatively simple situation for explanation purposes. It should be noted that the formulation of the estimators, presented in Equations 4-5 and 7-8, is a function with parent numbers for each population; in the next section, we provide the treatment for estimating the unknown parent numbers by additional usage of HS and/or PO pairs. The main symbols used in the current paper are summarized in {\bf Table \ref{symbols}}.

@@

Suppose that we have a set of two populations labelled 1 and 2 where random mating occurs within each population and that parents can move to the other population after the end of the reproductive season. Hereafter, without loss of generality, we consider the movement of parents in the direction from population 1 to population 2 and focus on the migration number or rate, and that is what we attempt to estimate. In this framework, considering two reproductive seasons (hereafter, we call them "the first year" and "the second year") is enough to estimate it. {\bf Figure \ref{cartoon}} presents a schematic representation of the kinship relationships and movements of parents.

Suppose there are $N_1$ parents in the population 1 at the beginning of the first year. Each parent produces offspring in which the number is governed by the parent's reproductive potential, denoted by $\lambda_{i,1}$ ($i=1,2,\ldots,N_1$). After the reproductive season, part of the parents would start to moving towards the population 2; $M$ survived migrants arrive in the population 2 so that there are $N_2$ parents in the population 2 at the beginning of the second year ($M < N_2$). Similar to the population 1, $N_2$ parents produce offspring in which the number is determined by $\lambda_{j,2}$ ($j=1,2,\ldots,N_2$). Theoretically speaking, the number of offspring for each parent is a random variable with mean $\lambda_{i,1}$ or $\lambda_{j,2}$.

@@

To estimate the migration number or rate, we utilize the number of HS and/or PO pairs observed in a sample. In the both populations, $n_{\rm O, 1}$ and $n_{\rm O, 2}$ offspring are randomly sampled. Additionally, $n_{\rm P, 1}$ and $n_{\rm P, 2}$ parents are randomly sampled immediately after the end of the reproductive season in the first and second year, respectively. For mathematical tractability, parents have to survive during the reproductive season; therefore, both mother and father of a certain offspring have a potential to be sampled. Assuming invasive sampling, sampled parents in the population 1 are never sampled in the population 2. It should be noted that al the four kinds of a sample (i.e., $n_{\rm P, 1}$, $n_{\rm P, 2}$, $n_{\rm O, 1}$ and $n_{\rm O, 2}$) are not always required for estimating migration number or rate; what kinds of a sample are needed depends on the situation, as explained later.

@@

@@

@@

\renewcommand{\arraystretch}{0.6}
\begin{table}[tb]
\begin{center}
   \caption[]{Table 1: The list of mathematical symbols employed in the main text}
    \textbf {}\\[-4mm]
    \begin{tabular}{llc} \hline
       & & \\
	$n_{\rm P,1}, n_{\rm P, 2}$			& Sample number of parents in the population 1 and 2\\ 
		                						& \\
	$n_{\rm O, 1}, n_{\rm O, 2}$			& Sample number of offspring in the population 1 and 2\\ 
		                						& \\
	$N_{1}, N_{2}$						& Number of parents in the population 1 and 2 when sampled offspring are born\\
		                						& \\
	$M$								& Number of survived migrants of parents from the population 1 to population 2\\
		                						& \\
	$r$								& Sex ratio\\
		                						& \\
	$\pi_{\rm PO,1}, \pi_{\rm PO,2}$		& Probability that a randomly selected pair (mother and offspring) \\
	                							& shares a parent-offspring relationship within the population 1 and 2\\
									& \\
	$\pi_{\rm PO, bet}$					& Probability that a randomly selected pair (parent and offspring in the population 1 and 2, \\
	                							& respectively) shares a parent-offspring relationship between the population 1 and 2\\
									& \\
	$\pi_{\rm HS, bet}$					& Probability that a randomly selected pair (two offspring in the population 1 and 2) \\
	                							& shares a half-sibling relationship between the population 1 and 2\\
					                			& \\
	$\lambda_{i,1}, \lambda_{j,2}$			& Expected number of surviving offspring of parent $i$ and $j$ at sampling in the population 1 and 2\\
		                						& \\
	$\lambda_{i, M}$					& Expected number of surviving offspring of migrant $i$ at sampling\\
		                						& \\
	$k_{i,1}, k_{j,2}$					& Number of surviving offspring born to parent $i$ and $j$ in the population 1 and 2\\ 
	                							& \\
	$H_{\rm PO, 1}, H_{\rm PO, 2}$		& Number of parent-offspring pairs observed in samples within the population 1 and 2\\ 
	                							& \\
	$H_{\rm PO, bet}$					& Number of parent-offspring pairs observed in samples between the population 1 and 2\\ 
	                							& \\
	$H_{\rm HS, bet}$					& Number of half-sibling pairs observed in samples between the population 1 and 2\\ 
	                							& \\              		
	\hline
    \end{tabular}
    \label{symbols} 
\end{center} 
\end{table}
\renewcommand{\arraystretch}{1}

\renewcommand{\arraystretch}{0.6}
\begin{table}[tb]
\begin{center}
   \caption[]{Table 2: Summary of proposed estimators for required parameters and kinship types}
    \textbf {}\\[-0mm]
    \begin{tabular}{ccccccccc} \hline
       	\\
	Estimator			& $N_1$ 		& $N_2$ 		& $n_{\rm O, 1}$	& $n_{\rm O, 2}$	& $n_{\rm P, 1}$	& $n_{\rm P, 2}$	& Required kinship type\\
	\\
	\hline
	\\
	$\widehat{M_1}$	& given 		& given 		& \checkmark		& \checkmark 		&				&				& HSP\\ 
		                						& \\
	$\widehat{M_2}$	& given		& given		& \checkmark		& 				&				& \checkmark 		& POP\\
		                						& \\
	$\widehat{M_3}$	& given		& given		& \checkmark		&  \checkmark		&				& \checkmark 		& HSP \& POP\\
		                						& \\
	$\widehat{M_4}$	& given		& estimated	& \checkmark		&  \checkmark		&				& \checkmark 		& HSP \& POP\\
		                						& \\
	$\widehat{M_5}$	& estimated	& estimated	& \checkmark		&  \checkmark		& \checkmark		& \checkmark 		& HSP \& POP\\
									& \\
	$\widehat{m_1}$	& ---			& given		& \checkmark		& \checkmark 		&				&				& HSP\\ 
									& \\
	$\widehat{m_2}$	& ---			& given		& \checkmark		& 				&				& \checkmark 		& POP\\					                			& \\
	$\widehat{m_3}$	& ---			& given 		& \checkmark		&  \checkmark		&				& \checkmark 		& HSP \& POP\\
		                						& \\
	$\widehat{m_4}$	& ---			& estimated	& \checkmark		&  \checkmark		& 				& \checkmark 		& HSP \& POP\\
		                						& \\
	\hline
    \end{tabular}
    \label{parameter} 
\end{center} 
\end{table}
\renewcommand{\arraystretch}{1}

@@

\section*{Reproductive potential}

\renewcommand{\theequation}{A\arabic{equation}}
Here, we introduce the concept of the reproductive potential of parent $i$ or $j$, which is defined as the expected number of surviving offspring at sampling time, denoted by $\lambda_{i,1}$ or $\lambda_{j,2}$. The reproductive potential is determined by several factors, including the parent’s age, weight, residence time on the spawning ground. It should be noted that the magnitude of this parameter ($\lambda_{i,1}$ or $\lambda_{i,2}$) includes information about the survival rate of the offspring, the number of days after egg hatching, and the egg number; this implies that the parameter reflects the sample timing. It should also be noted that the modeling framework does not depend on whether the reproductive potential is heritable or not.

@@

\renewcommand{\theequation}{B\arabic{equation}}
